---
title: "Data Manipulation"
output: github_document
---

This document will show how to *manipulate* data. 

Importing the two datasets pups and litters that we're going to manipulate. 

```{r setup, echo = FALSE, message = FALSE}
library(tidyverse)

```

```{r}
litters_df = 
  read_csv("data/FAS_litters.csv", na = c("NA", "","."))

litters_df = janitor::clean_names(litters_df)

pups_df = 
  read_csv("data/FAS_pups.csv", na = c("NA", "","."))

pups_df = janitor::clean_names(pups_df)
```

## `select`

Use `select()` to select variables by naming the data frame (litters_df) and different variables you want to select. 

```{r}
select(litters_df, group, litter_number, gd0_weight)
```

You can also select a range of columns - group to gd18_weight
```{r}
select(litters_df, group:gd18_weight)
```

You can also remove columns with the - sign.
```{r}
select(litters_df, -pups_survive)
```

You can also remove a range of columns. 
```{r}
select(litters_df, -(group:gd18_weight))
```

```{r}
select(litters_df, starts_with("gd"))

select(litters_df, contains("pups"))
```

You can also rename columns and rearrange columns (last 2 lines). Can use relocate in place of select to rearrange columns. 
```{r}
select(litters_df, GROUP = group)

rename(litters_df, GROUP = group)

select(litters_df, litter_number, gd0_weight, everything())

relocate(litters_df, litter_number, gd0_weight)
```

```{r learning_assessment}
select(pups_df, litter_number, sex, pd_ears)
```

## `filter`

## `mutate`

## `arrange`

## `PIPING`