---
title: "Data Import document"
output: github_document
---

```{r setup, echo = FALSE, message = FALSE}
library(tidyverse)
```

This document will show how to import data. 

The clean_names function in janitor package cleans the name of the variables to lower case and replaces space with dash.
```{r}
litters_df = read_csv("data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```

## Look at data set

```{r}
litters_df
head(litters_df)
tail(litters_df)
```

The view() function will open the entire data frame in a new window.
eval = FALSE so the data set doesn't show up in the knitted document.

```{r eval = FALSE}
#view will open the entire data frame in a new window 
#eval = FALSE so the data set doesn't show up in the knitted document
view(litters_df)

```

## Learning Assessment
Importing the FAS_pups dataset. 

Use relative path 
```{r}
pups_df = read_csv("./data/FAS_pups.csv")
pups_df = janitor::clean_names(pups_df)
```

Use absolute path 
```{r absolute_path}
#pups_df = read_csv("/Users/claudiahuynh/Desktop/BIST8105/data_wrangling_i/data")

```

## Look at read_csv options 

col_names = FALSE function generates column names as X1, X2, X3
skip = 1 skips the first row which was the original col names 
```{r}
litters_df = 
  read_csv(
    file = "data/FAS_litters.csv",
    col_names = FALSE,
    skip = 1
  )
```

What about missing data?
Use the na = c() function to treat missing data presented in the forms of Na, space or . as NA.

```{r}
litters_df = 
  read_csv(
    file = "data/FAS_litters.csv",
    na = c("NA", "", "."),
  )

```

What if we code "group" as a factor variable?
Factor variables are useful when we know that the column has different levels. This section does column parsing using the col_types = col() function.  

```{r}
litters_df = 
    read_csv(
      file = "data/FAS_litters.csv",
      col_types = cols(
        Group = col_factor()
      )
    )
```


