---
title: "Data Manipulation"
output: github_document
---

This document will show how to *manipulate* data. 

Importing the two datasets pups and litters that we're going to manipulate. 

```{r setup, echo = FALSE, message = FALSE}
library(tidyverse)

```

```{r}
litters_df = 
  read_csv("data/FAS_litters.csv", na = c("NA", "","."))

litters_df = janitor::clean_names(litters_df)

pups_df = 
  read_csv("data/FAS_pups.csv", na = c("NA", "","."))

pups_df = janitor::clean_names(pups_df)
```

## `select`

Use `select()` to select variables by naming the data frame (litters_df) and different variables you want to select. 

```{r}
select(litters_df, group, litter_number, gd0_weight)
```

You can also select a range of columns - group to gd18_weight
```{r}
select(litters_df, group:gd18_weight)
```

You can also remove columns with the - sign.
```{r}
select(litters_df, -pups_survive)
```

You can also remove a range of columns. 
```{r}
select(litters_df, -(group:gd18_weight))
```

```{r}
select(litters_df, starts_with("gd"))

select(litters_df, contains("pups"))
```

You can also rename columns and rearrange columns (last 2 lines). Can use relocate in place of select to rearrange columns. 
```{r}
select(litters_df, GROUP = group)

rename(litters_df, GROUP = group)

select(litters_df, litter_number, gd0_weight, everything())

relocate(litters_df, litter_number, gd0_weight)
```

```{r learning_assessment}
select(pups_df, litter_number, sex, pd_ears)
```

## `filter`
You can filter values using ==
```{r}
filter(litters_df, gd_of_birth == 20)
filter(litters_df, gd_of_birth == 19)
```

```{r}
filter(litters_df, gd_of_birth == 20)
filter(litters_df, gd_of_birth == 19)
```

You can also do inequalities 

```{r}
filter(litters_df, pups_born_alive >8)
filter(litters_df, pups_born_alive >=8)
```

You can also do not equal to using !=

```{r}
filter(litters_df, pups_born_alive !=9)

```

You can also filter character columns. Filter multiple with %in%
```{r}
filter(litters_df, group == "Low8")
filter(litters_df, group %in% c("Low7", "Low8"))
```

You can also filter multiple rows
```{r}
filter(litters_df, group %in% c("Low7", "Low8"), pups_born_alive ==8)
```

You can also get rid of missing values 
```{r}
drop_na(litters_df)

drop_na(litters_df, gd0_weight)
```

Learning assessment 2
```{r la2}
filter(pups_df, sex == 1)
filter(pups_df, pd_walk <11, sex == 2)
```
## `mutate`
```{r}
mutate(litters_df, wt_gain = gd18_weight - gd0_weight)

mutate(litters_df, sq_pups = pups_born_alive^2)

mutate(litters_df, group = str_to_lower(group))
```

```{r}
mutate(
  litters_df,
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group)
)
```

## `arrange`
Reorder the rows to ascending or descending order. You can also arrange multiple rows. For example, arrange pups_born_alive first then gd0 weight. 

```{r}
arrange(litters_df, gd0_weight)

arrange(litters_df, desc(gd0_weight))

arrange(litters_df, pups_born_alive, gd0_weight)
```

## `PIPING`!!!!
